---
title: 'Statistical report: Covid-19 vs sglt2i/glp1a/dpp4i'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data sources

### SCB

Län (Stockholm), civil status and family type are from RTB as of 2019-12-31. 
Age, sex, country of birth and number of children is also from 2019-12-31.
Disposable income and education are from LISA 2018. Age is from 2018 in the 
negative control analyses (see below). 

### Socialstyrelsen

Comorbidities and covid-19 outcome are derived from the National Patient Registry 
(in-patient or out-patient) from 1997 (1998 Skåne) 
and treatments from the Dispensed Drug Registry. 

Note that data after 2018 is preliminary for the National Patient Registry and 
Cause of Death Registry. From Socialstyrelsens homepage:
"Patientregistret kommer göras tillgängligt för utlämnanden för forskningsprojekt
och statistikbeställningar redan en eller två månader efter referensmånaden. De
data som görs tillgängliga kommer delvis vara ofullständiga, framför allt
kommer det med hög sannolikhet att saknas diagnoskoder för delar av
materialet. Regionala skillnader avseende kvalitet kommer därför att finnas i de
data som görs tillgängliga. Fullständiga data i patientregistret beräknas finnas tre
månader efter rapporterad period.
Dödsorsaksintyg för dödsfall som är relaterade till covid-19 kommer kodas med
förtur, så att data om sådana dödsfall kommer vara tillgängliga cirka fyra veckor
efter dödsfallet"

Information from handläggare at Socialstyrelsen via e-mail: 
"patientregistret slutenvård år 2020 kan det finnas pågående vårdtillfällen 
(dvs. vårdtillfällen utan  utskrivningsdatum). Jag vet inte hur ni hanterar 
dessa, men det kan vara bra för dig att veta. Eftersom diagnos ofta fastställt 
vid utskrivning kan diagnoserna på dessa vårdtillfällen vara lite ”skakiga”. 

## Time period

Index date is defined as `r global_indexdtm` (first covid-19 case in Sweden end Jan). 
End of follow-up is defined as `r global_enddtm`, 
decision based on visual inspection of patterns in NPR and Cause of Death Registry data.

## Population 

The population received from Socialstyrelsens consisted of persons with any of the following: 
hypertension, diabetes, ischemic heart disease, heart failure, kidney disease, obesity or af 
(ICD codes in Table \ref{tab:outcomcodes}) defined 
from the NPR (in-patient or out-patient) from 1997 (1998 Skåne) and alive 2020-01-01. 

\clearpage
\newpage

```{r flow}
mykable(flow, caption = "Flowchart", scale_down = F)
```

The median age (IQR) is `r pop %>% summarise(med = dF(median(scb_age), dig = 1),
                                             q1 = dF(quantile(scb_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(scb_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pop %>% count(scb_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(scb_sex == "Female") %>%
  pull(perc)`% females.                                   

## Created variables 

```{r outcomcodes}
mykable(metaout, caption = "Comorbidities and outcomes from NPR")
```

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

Time on treatment is calculated as time from first (from 1 July 2005, start of DDR) 
to last  + 3 months (prior to `r global_indexdtm`) dispensed prescription for the respective treatment.

### Outcome Covid-19

The outcome is defined as hospitalized (in-patient) confirmed covid-19 (U071) as main 
diagnosis (HDIA) in the National Patient Registry OR as confirmed covid-19 (U071)
as underlying cause of death (ULORSAK) in Cause of Death Registry. 

```{r, child = "../src/deskstatcovid.Rmd"}

```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

There was missing data for some of the scb variables (not sex, age, number of children). 
Patients with missing data are excluded from all analysis. 

## Propensity scores and matching

A propensity score for treatment with sglt2i, glp1a and dpp4i respectively was estimated for 
each patient with logistic regression with variables indicated in Table \ref{tab:tab1allsglt2match}, partly in 
all patients, partly in the patients with covid-19. Continuous variables (age) 
were modelled using cubic splines with 4 dfs. 

1:1 matching without replacement [@match] was thereafter performed where the propensity 
score was allowed to differ by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

1:1 has deemed the best option when taking into account the number of patients retained.

## Baseline characteristics

```{r, child = "../src/tab1match.Rmd"}

```

\clearpage
\newpage

## Covid-19

The association between sglt2i, glp1a and dpp4i separately and Covid-19 was investigated 
using the modified Poisson regression approach [@gee], i.e. using 
Generalized Estimating Equations (GEE) models with a Poisson distribution and a robust error variance, 
for the following models: 

- Crude
- Adjusted for the individual variables indicated in Table \ref{tab:tab1allsglt2match} 
- In the matched cohort, so therefore adjusted for the same variables as above

In the model for the matched cohort the matched pairs were modelled using an exchangeable correlation structure. 

The association in predefined subgroups were investigated using interaction effects.

### Consistency analysis

#### Competing risk

The analyses were also performed using a sub-distributional hazards model [@fg] 
were death was treated as a competing event.

For the main analysis censoring is not incorporated in the model since this also 
introduces time as a factor which is probably more due to spread of disease. 

#### Incident user 

The analyses were also performed were patients collecting prescriptions within 1 year 
prior to the 5 month period defining if the patient is a treatment user were excluded from the analysis, 
thus only including incident users [@iu]. Note that also patients in the non-treatment group can 
be excluded if they have a dispensed prescription during this period. 

\clearpage
\newpage

```{r, child = "../src/outcovidtab.Rmd"}

```

\clearpage
\newpage

## All-cause mortality after Covid-19

```{r, child = "../src/deskstatdeaths.Rmd"}

```

The time to all-casues mortality is presented in 1-KM curves for the matched population.

The association between treatments and all-cause mortality after Covid-19 was 
investigated using cox regression for the following models: 

- Crude
- Adjusted for the individual variables indicated in Table \ref{tab:tab1allsglt2match} 
- In the matched cohort, so therefore adjusted for the same variables as above

In the model for the matched cohort the matched pairs were incorporated in the 
model with a strata term. 

The association in predefined subgroups were investigated using interaction effects.

Only patients with Covid-19 were included in the models. 

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the matched models. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> No action deemed necessary. 

\clearpage
\newpage

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/covid19sglt2i. On publication 
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
